Robust receding horizon control for linear systems with state and input dependent disturbances

### 2015-06-21 example.tex
Added review version.

%!TEX root = draft_v2.tex
\section{Complexity}\label{sec:complexity}
We next discuss bounds on the computational complexity of the algorithm outlined in Section~\ref{sec:line:search}.
%The computational complexity of the considered algorithm can easily be bounded. 
%In order to give implementation-free bounds  these are expressed 
%These are given in terms of floating point operations
%(flops). 
For a given active set at most $2N$ mpQPs must be solved, and this is done by decomposing the
problem as stated in section~\ref{sec:gen:mpQP}.  For each mpQP the main computational effort is to
determine $\Sigma$ in~\eqref{seq:sigma}. In the case of~\eqref{eq:sigma:normal} this requires a QR
decomposition with $2(mn^2-n^3/3)$ floating point operations (flops), where $m$ is the number of active constraints and $n$ the number of
optimisation variables, plus inversion of a triangular matrix with $(m^3-m)/6$ flops and several matrix
multiplications with $mpn$ flops (for the multiplication of a $m\times n$ matrix with a $n\times p$
matrix). The computationally more expensive case of~\eqref{eq:sigma:degen} requires a QR decomposition,
inversion of a definite matrix $W$ using Cholesky decomposition with $n^3/3$ flops, inversion of a triangular
matrix, several matrix multiplications and additions with $mn$ flops, and inversion of a definite matrix
$\Delta$. In the worst case this adds up to at most $11n^3+n^2-n/3$ flops to obtain $\Sigma$. The solution to
the problem is then the result of a matrix multiplication, and, in order to be able to solve the subsequent
mpQP, the matrix $W_\theta$ and the vector $c_\theta$ also need to be computed. Hence the overall flop count
in the worst case adds up to $11n^3+n^2+2/3n+2n^2 n_\theta$, where $n_\theta$ denotes the dimension of the
parameter~$\theta$. 
%At each stage two mpQPs have to be solved recursively, so that 
In the worst case the
recursion for the entire horizon therefore requires $N(2 n_x+5 n_x^2+101 n_x^3+8 n_x^2 (n_u+n_x))$ flops.

>>>>

%!TEX root = draft_v2.tex
\section{Complexity}\label{sec:complexity}
We next discuss bounds on the computational complexity of the algorithm outlined in Section~\ref{sec:line:search}.
For a given active set at most $2N$ mpQPs must be solved, and this is done by decomposing the
problem as stated in section~\ref{sec:gen:mpQP}.  For each mpQP the main computational effort is to
determine $\Sigma$ in~\eqref{seq:sigma}. In the case of~\eqref{eq:sigma:normal} this requires a QR
decomposition with $2(mn^2-n^3/3)$ floating point operations (flops, see e.g.~\cite{Golub:1996}), 
where $m$ is the number of active constraints and $n$ the number of
optimisation variables, plus inversion of a triangular matrix with $(m^3-m)/6$ flops and several matrix
multiplications with $mpn$ flops (for the multiplication of a $m\times n$ matrix with a $n\times p$
matrix). The computationally more expensive case of~\eqref{eq:sigma:degen} requires a QR decomposition,
inversion of a definite matrix $W$ using Cholesky decomposition with $n^3/3$ flops, inversion of a triangular
matrix, several matrix multiplications and additions with $mn$ flops, and inversion of a definite matrix
$\Delta$. In the worst case this adds up to at most $11n^3+n^2-n/3$ flops to obtain $\Sigma$. The solution to
the problem is then the result of a matrix multiplication, and, in order to be able to solve the subsequent
mpQP, the matrix $W_\theta$ and the vector $c_\theta$ also need to be computed. Hence the overall flop count
in the worst case adds up to $11n^3+n^2+2/3n+2n^2 n_\theta$, where $n_\theta$ denotes the dimension of the
parameter~$\theta$. 
%At each stage two mpQPs have to be solved recursively, so that 
In the worst case the
recursion for the entire horizon therefore requires $N(2 n_x+5 n_x^2+101 n_x^3+8 n_x^2 (n_u+n_x))$ flops.

### 2015-06-21 post_review.tex
Added review version.

\section{Controllable sets}\label{sec:preliminaries}
In this section we summarise results on controllable sets that allow the constraints of the problem~\eqref{eq:the:problem} to be to formulated
in a convenient way. The following result shows that the MRPI set (\ref{eq:definition:MRPI:set}) associated
with the pointwise polytopic set~\eqref{eq:PWA:def:disturbance} is polytopic if $H$ is elementwise convex and
piecewise affine. For a proof and detailed discussion of this result we refer the reader
to~\cite{Schaich:2015}.
%
\begin{lem}\label{thm:MRPI:set:is:polytopic}
Let $\mathcal X \subseteq \{x:\Gamma x\leq{\bf{1}}\wedge-\Gamma x\leq{\bf{1}}\}$ be polyhedral, and the pair $(A+BK,\Gamma)$ be
observable, furthermore let $\mathcal W(x,Kx)$ be defined as in~\eqref{eq:PWA:def:disturbance} with a elementwise convex piecewise
affine function $H(x,Kx)$ which is bounded for all finite $x$. Then the MRPI set $\mathcal X^\infty$ is a polytope, i.e.
it has the representation 
\begin{equation}\label{eq:halfspace:representation:MRPI}
	\mathcal X^\infty=\{x:\Lambda_i x\leq \lambda_i, \forall\, i\in\mathcal I^\infty\}.
\end{equation}
\end{lem}
%
We need to emphasise the following 
\begin{rem}\label{rem:vertex:representation}
Note that an essential assumption in Lemma~\ref{thm:MRPI:set:is:polytopic} is that $\mathcal W(x,u)$ is a polytope
for all finite $(x,u)$, i.e. it has a convex hull representation $\mathcal W(x,u) = \conv_k\{w_k(x,u)\}, k\in\{1,\dots,r\}$ where each 
vertex, $w_k(x,u)$, is a piecewise affine function of $(x,u)$, i.e.
\begin{multline}
	w_k(x,u) = \max\{W_{k,1}^x x + W_{k,1}^u u + w_{k,1}, \\
        W_{k,2}^x x + W_{k,2}^u u + w_{k,2},\dots\}
\end{multline}
where the maximisation is performed elementwise.
\end{rem}
%
We now derive the representation of $\mathcal Z_m$ and $\mathcal X_m$ as defined in~\eqref{eq:definition:stage:constraints}.
According to Lemma~\ref{thm:MRPI:set:is:polytopic} the MRPI set $\mathcal X^\infty$ is polytopic and has the 
representation~\eqref{eq:halfspace:representation:MRPI}. Therefore the recursion~\eqref{eq:definition:mixed:stage:constraint}
defines $\mathcal Z_1$ as
\begin{equation}\label{eq:derivation:of:mpLP:stage:constraints}
\begin{aligned}
	\mathcal Z_1 =& (\mathbb R^n\times\mathcal U)\cap 
\begin{aligned}[t] \Bigl\{(x,u): \Lambda_{0,i}(Ax&+Bu+Dw)\leq \lambda_{0,i} \\
& \forall w\in\mathcal W(x,u),\, i\in\mathcal I_0 \Bigr\} \end{aligned}
\\
	=&\begin{aligned}[t] (\mathbb R^n & \times \mathcal U) \cap\Bigl\{ (x,u): \\
& \max_{w\in\mathcal W(x,u)}\Lambda_{0,i}(Ax+Bu+Dw)\leq\lambda_{0,i},\, \forall i\in\mathcal I_0 \Bigr\} \end{aligned}
\\
	=&\begin{aligned}[t] & (\mathbb R^n \times\mathcal U)\cap \Bigl\{ (x,u): \\
& \!\!\!\!\!\!
\Lambda_{0,i}(Ax+Bu)+ \!\!\underbrace{\max_{j\in\{1,\dots,r\}}\!\Lambda_{0,i} Dw_j(x,u)}_{(i)}\leq\lambda_i, \, \forall i\in\mathcal I_0
	\Bigr\} \end{aligned}
\end{aligned}
\end{equation}
Since the maximisation $(i)$ in \eqref{eq:derivation:of:mpLP:stage:constraints} is a multi-parametric
linear program, its solution is piecewise affine in $(x,u)$. 
Therefore the set $\mathcal Z_1$ has a representation
$\mathcal Z_1=\{(x,u):\Xi^x_{1,i}x+\Xi^u_{1,i}u\leq\xi_{1,i},i\in\mathcal J_1\}$. The projection~$\mathcal X_1$ 
of $\mathcal Z_1$ onto~$\mathbb R^n$ is therefore has the representation~$\mathcal X_1 = \{x:\Lambda_{1,i} x\leq\lambda_{1,i},i\in\mathcal I_1\}$.
From here the determination of $\mathcal Z_k$ and $\mathcal X_k$ follows recursively. We are now able to formulate 
all constraints in~\eqref{eq:the:problem} as inequalities. In the next section we discuss a method of solving
min-max programs such as~\eqref{eq:the:problem} efficiently.

@@

& + m^TWM_\theta+\frac{1}{2}m^T Wm + c^Tm\\

@@

We recursively define all primal and dual variables:
\begin{equation}\label{eq:trajectory:affine:representation}
	\begin{split}
		u_N(x_N) =& \underbrace{K_{x,N}^x}_{\Psi_N^u} x_N + \underbrace{k_N^u}_{\psi_N^u}\\
		\nu_N(x_N,\beta_N) &= \underbrace{L_{x,N}^x}_{\Psi_N^\nu} x_N + \underbrace{L_{\beta,N}^\nu}_{\Phi_N^\nu} \beta_N + 
		\underbrace{l_N^\nu}_{\psi_N^u}\\
		&\vdots \\
		x_0(x_N) =& \underbrace{M_{z,1}^{x^+}\left(\begin{array}{c} \Psi_2^x \\ \Psi_1^u \end{array}\right)}_{\Psi_1^x}x_N + 
		\underbrace{M_{z,1}^{x^+}\left(\begin{array}{c} \psi_2^x \\ \psi_1^u \end{array}\right) + m_1^{x^+}}_{\psi^x_1}\\
		\eta_1(x_N,\beta_N) =& \underbrace{\left(N_{z,1}^{\eta}\left(\begin{array}{c} \Psi_2^x \\ \Psi_1^u \end{array}\right) + 
		N_{\hat\beta,1}^\eta \Psi_1^{\hat\zeta}\right)}_{\Psi_1^\eta}x_N + \underbrace{N_{\hat\beta,1}^\eta \Phi_1^{\hat\zeta}}_{\Phi_1^\eta} \beta_N\\
		&+\underbrace{N_{z,1}^{\eta}\left(\begin{array}{c} \psi_2^x \\ \psi_1^u \end{array}\right) + 
		N_{\hat\beta,1}^\eta \psi_1^{\hat\zeta} + n_1^\eta}_{\psi_1^\eta}\\
		&\vdots
	\end{split}
\end{equation}
Notice that $\Phi^{(\cdot)}_m$ is only non-trivial if the associated sub-problem is degenerate. The
representation~\eqref{eq:trajectory:affine:representation} is valid as long as the constraint set remains
unchanged, i.e.\ the inequalities that hold with equality do not change. Starting from the known set of active
constraints at $\chi_0$ we perform a line search towards $\chi_e$, i.e. we vary
$x_N(\alpha)=\chi_0+ \alpha(\chi_e-x_0)$ for $0\leq\alpha\leq1$. We
use~\eqref{eq:trajectory:affine:representation} to check all inactive constraints with the primal variables
$x_m(x_N(\alpha)),\, u_m(x_N(\alpha))$ and~$w_m(x_N(\alpha))$, and all dual variables corresponding to
inequality constraints to check whether constraints need to be deactivated. That is, we search for the
smallest $\alpha$ such that $\Xi_{m,i}^x x_m(x_N(\alpha))+\Xi_{m,i}^u u_m(x_N(\alpha))=\xi_{m,i}$ or
$G_iw_m(x_N(\alpha))=1$ or $\eta_{m,i}(x_N(\alpha),\beta_N)=0$ for some $m$ and $i$, and update the set of active
constraints accordingly, i.e. if an additional constraint is activated it is introduced to the equality
constraint mpQP recursion and the updated problems are solved again recursively, if a constraint is
deactivated because the smallest $\alpha$ corresponds to the zero-crossing of its dual variable then it is
removed and the
mpQP recursion is solved again. The updated active set is the new known active set at $\chi_0 = x_N(\alpha)$
and the line-search continues until the state~$\chi_e$ is in the same active set as $\chi_N(\alpha)$ where the
line-search terminates.  To initialise we start at the origin $\chi_0=0$, since we know that no constraints
are active.

>>>>

\section{Controllable sets}\label{sec:preliminaries}
In this section we summarise results on controllable sets that allow the constraints of the problem~\eqref{eq:the:problem} to be to formulated
in a convenient way. The following result shows that the MRPI set (\ref{eq:definition:MRPI:set}) associated
with the pointwise polytopic set~\eqref{eq:PWA:def:disturbance} is polytopic if $H$ is elementwise convex and
piecewise affine. For a proof and detailed discussion of this result we refer the reader
to~\cite{Schaich:2015}.
%
\begin{lem}\label{thm:MRPI:set:is:polytopic}
Let $\mathcal X \subseteq \{x:\Gamma x\leq{\bf{1}}\wedge-\Gamma x\leq{\bf{1}}\}$ be polyhedral, and the pair $(A+BK,\Gamma)$ be
observable, furthermore let $\mathcal W(x,Kx)$ be defined as in~\eqref{eq:PWA:def:disturbance} with a elementwise convex piecewise
affine function $H(x,Kx)$ which is bounded for all finite $x$. Then the MRPI set $\mathcal X^\infty$ is a polytope, i.e.
it has the representation 
\begin{equation}\label{eq:halfspace:representation:MRPI}
	\mathcal X^\infty=\{x:\Lambda_i x\leq \lambda_i, \forall\, i\in\mathcal I^\infty\}.
\end{equation}
\end{lem}
%
We need to emphasise the following 
\begin{rem}\label{rem:vertex:representation}
Note that an essential assumption in Lemma~\ref{thm:MRPI:set:is:polytopic} is that $\mathcal W(x,u)$ is a polytope
for all finite $(x,u)$, i.e. it has a convex hull representation $\mathcal W(x,u) = \conv_k\{w_k(x,u)\}, k\in\{1,\dots,r\}$ where each 
vertex, $w_k(x,u)$, is a piecewise affine function of $(x,u)$, i.e.
\begin{multline}
	w_k(x,u) = \max\{W_{k,1}^x x + W_{k,1}^u u + w_{k,1}, \\
        W_{k,2}^x x + W_{k,2}^u u + w_{k,2},\dots\}
\end{multline}
where the maximisation is performed elementwise.
\end{rem}
%
We now derive the representation of $\mathcal Z_m$ and $\mathcal X_m$ as defined in~\eqref{eq:definition:stage:constraints}.
According to Lemma~\ref{thm:MRPI:set:is:polytopic} the MRPI set $\mathcal X^\infty$ is polytopic and has the 
representation~\eqref{eq:halfspace:representation:MRPI}. Therefore the recursion~\eqref{eq:definition:mixed:stage:constraint}
defines $\mathcal Z_1$ as
\begin{equation}\label{eq:derivation:of:mpLP:stage:constraints}
\begin{aligned}
	\mathcal Z_1 =& (\mathbb R^n\times\mathcal U)\cap 
\begin{aligned}[t] \Bigl\{(x,u): \Lambda_{0,i}(Ax&+Bu+Dw)\leq \lambda_{0,i} \\
& \forall w\in\mathcal W(x,u),\, i\in\mathcal I_0 \Bigr\} \end{aligned}
\\
	=&\begin{aligned}[t] (\mathbb R^n & \times \mathcal U) \cap\Bigl\{ (x,u): \\
& \max_{w\in\mathcal W(x,u)}\Lambda_{0,i}(Ax+Bu+Dw)\leq\lambda_{0,i},\, \forall i\in\mathcal I_0 \Bigr\} \end{aligned}
\\
	=&\begin{aligned}[t] & (\mathbb R^n \times\mathcal U)\cap \Bigl\{ (x,u): \\
& \!\!\!\!\!\!
\Lambda_{0,i}(Ax+Bu)+ \!\!\underbrace{\max_{j\in\{1,\dots,r\}}\!\Lambda_{0,i} Dw_j(x,u)}_{(i)}\leq\lambda_i, \, \forall i\in\mathcal I_0
	\Bigr\} \end{aligned}
\end{aligned}
\end{equation}
Since the maximisation $(i)$ in \eqref{eq:derivation:of:mpLP:stage:constraints} is a multi-parametric
linear program, its solution is piecewise affine in $(x,u)$. 
Therefore the set $\mathcal Z_1$ has a representation
$\mathcal Z_1=\{(x,u):\Xi^x_{1,i}x+\Xi^u_{1,i}u\leq\xi_{1,i},i\in\mathcal J_1\}$. The projection~$\mathcal X_1$ 
of $\mathcal Z_1$ onto~$\mathbb R^n$ therefore has the representation~$\mathcal X_1 = \{x:\Lambda_{1,i} x\leq\lambda_{1,i},i\in\mathcal I_1\}$.
From here the determination of $\mathcal Z_k$ and $\mathcal X_k$ follows recursively. We are now able to formulate 
all constraints in~\eqref{eq:the:problem} as inequalities. In the next section we discuss a method of solving
min-max programs such as~\eqref{eq:the:problem} efficiently.

@@

& + m^TWM_\theta \theta+\frac{1}{2}m^T Wm + c^Tm\\

@@

We recursively define all primal and dual variables:
\begin{equation}\label{eq:trajectory:affine:representation}
	\begin{split}
		u_N(x_N) =& \underbrace{K_{x,N}^u}_{\Psi_N^u} x_N + \underbrace{k_N^u}_{\psi_N^u}\\
		\nu_N(x_N,\beta_N) &= \underbrace{L_{x,N}^\nu}_{\Psi_N^\nu} x_N + \underbrace{L_{\beta,N}^\nu}_{\Phi_N^\nu} \beta_N + 
		\underbrace{l_N^\nu}_{\psi_N^u}\\
		&\vdots \\
		x_0(x_N) =& \underbrace{M_{z,1}^{x^+}\left(\begin{array}{c} \Psi_2^x \\ \Psi_1^u \end{array}\right)}_{\Psi_1^x}x_N + 
		\underbrace{M_{z,1}^{x^+}\left(\begin{array}{c} \psi_2^x \\ \psi_1^u \end{array}\right) + m_1^{x^+}}_{\psi^x_1}\\
		\eta_1(x_N,\beta_N) =& \underbrace{\left(N_{z,1}^{\eta}\left(\begin{array}{c} \Psi_2^x \\ \Psi_1^u \end{array}\right) + 
		N_{\hat\beta,1}^\eta \Psi_1^{\hat\zeta}\right)}_{\Psi_1^\eta}x_N + \underbrace{N_{\hat\beta,1}^\eta \Phi_1^{\hat\zeta}}_{\Phi_1^\eta} \beta_N\\
		&+\underbrace{N_{z,1}^{\eta}\left(\begin{array}{c} \psi_2^x \\ \psi_1^u \end{array}\right) + 
		N_{\hat\beta,1}^\eta \psi_1^{\hat\zeta} + n_1^\eta}_{\psi_1^\eta}\\
		&\vdots
	\end{split}
\end{equation}
Notice that $\Phi^{(\cdot)}_m$ is only non-trivial if the associated sub-problem is degenerate. The
representation~\eqref{eq:trajectory:affine:representation} is valid as long as the constraint set remains
unchanged, i.e.\ the inequalities that hold with equality do not change. Starting from the known set of active
constraints at $\chi_0$ we perform a line search towards $\chi_e$, i.e. we vary
$x_N(\alpha)=\chi_0+ \alpha(\chi_e-x_0)$ for $0\leq\alpha\leq1$. We
use~\eqref{eq:trajectory:affine:representation} to check all inactive constraints with the primal variables
$x_m(x_N(\alpha)),\, u_m(x_N(\alpha))$ and~$w_m(x_N(\alpha))$, and all dual variables corresponding to
inequality constraints to check whether constraints need to be deactivated. That is, we search for the
smallest $\alpha$ such that $\Xi_{m,i}^x x_m(x_N(\alpha))+\Xi_{m,i}^u u_m(x_N(\alpha))=\xi_{m,i}$ or
$G_iw_m(x_N(\alpha))=1$ or $\eta_{m,i}(x_N(\alpha),\beta_N)=0$ for some $m$ and $i$, and update the set of active
constraints accordingly, i.e. if an additional constraint is activated it is introduced to the equality
constraint mpQP recursion and the updated problems are solved again recursively, if a constraint is
deactivated because the smallest $\alpha$ corresponds to the zero-crossing of its dual variable then it is
removed and the
mpQP recursion is solved again. The updated active set is the new known active set at $\chi_0 = x_N(\alpha)$
and the line-search continues until the state~$\chi_e$ is in the same active set as $\chi_N(\alpha)$ where the
line-search terminates.  To initialise we start at the origin $\chi_0=0$, since we know that no constraints
are active.
